---
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Before you do anything else

If you only want to update packages (not R, RStudio etc.) then:

0) This is just a generic recomendation when working in RStudio:
  * Go to Tools -> Global option ... -> General
    * Make sure that "Restore .RData into workspace at startup" is **not selected**
    * Set "Save workspace to .RData on exit" to "**Never**"
  * Go to Tools -> Global option ... -> R Markdown
    * Make sure that "Show output inline for all R Markdown documents" is **not selected**
1) Make sure you have only one RStudio open and that it is a "fresh" session
2) Do as described on [gitlab](https://gitlab.hafogvatn.is/-/snippets/3):

```{r eval = FALSE}
options(repos = c(CRAN = "https://cran.hafro.is"))
update.packages(ask=FALSE, checkBuilt = TRUE,lib=Sys.getenv("R_LIBS_USER"))
```

# The quick installation guide

The most common usage of the {xe} packages is to serve as a basis for the "smxapp", a quality control shiny app that is used during the conduction of the Icelandic bottom trawl surveys. In order to provide also a historical comparisons older archieved survey data must also be available. These reside (as of 2022) in the {mardata} package. Since these data are not publicly available as of yet, that package is not distributed on github but rather internally ... To install it do:

```{r, eval = FALSE}
remotes::install_local("R:/R/Pakkar/mardata", force = TRUE)
```

**Note**: This assumes that you are connected to MFRI instute network, either at the office or via VPN. Hence it is strongly suggested that you install this package prior to leaving on a cruise.

The purpose of the {xe} package is to create a connection and some convenient functions to the MRI Oracle XE-database via R. Unlike the mar-database the XE-database resides on personal computers and is what the software [hafvog](https://heima.hafro.is/~darri/hafvog_vefur) communicates with. The {xe} packages stands on its own and can be used for ..., further details are provide [here](https://heima.hafro.is/~einarhj/xe/articles/overview_xe.html). To install it do:

```{r, eval = FALSE}
remotes::install_github("fishvice/xe",  dependencies = FALSE, args='--no-multiarch')
```

**As of 2022-02-22 the version of Hafvog is 4.1.8, including the setup (stillingar) and support tables (stoðtöflur).** Since the XE database is run on Oracle 11g one can not not use dbplyr versions >2.0. Hence we need to install older versions:

```{r, eval = FALSE}
remotes::install_github("tidyverse/dbplyr@v1.4.4", force = TRUE)
```

**Side effects**: Installing an earlier version of {dbplyr} means that one can not load the tidyverse package only its child packages (dplyr, ggplot2, tidyr, tibble, ....). It also means that if you want to access data on the main Oracle database via {mar} one has to reinstall the latest version of {dbplyr}. By  `install.packages("dbplyr"}` one overwrites {dbplyr} version 1.4.4 with the latest one available on cran.


# The smxapp

A template for the smxapp is available via the {xe} package. To access the template one does as follows within RStudio:

```
File -> New File -> R Markdown -> From template -> smx dashboard
```

## 2022 SMB testfile

A test file is availble on hafro ftp-site, you can download it directly via:

```{r, eval = FALSE}
download.file("ftp.hafro.is/pub/data/TB1-2022.zip", destfile = "TB1-2022.zip")
```

.. or you can use Filezilla and the non-reproducable mouse-click method.
